\documentclass[12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load in packages
\usepackage{amsmath}
\usepackage{float}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{center}
\Large Import Tariffs and Export Duties

\medskip

\normalsize Elements of Microeconomics (discussion section 1)

\medskip

\small Rudy Kretschmar
\end{center}
\section*{Question 1}
Narnia is a large land that is far, far away. During the reign of the White Witch, 
the land produces and consumes Turkish Delight. The world price of Turkish Delight 
is \$2 per bag.

Narnia's demand and supply for Turkish Delight are governed by the following equations:
\begin{align*}
    Q_D &= 8 - P \\
    Q_S &= P
\end{align*}

where \(P\) is measured in dollars per bag and \(Q\) is measured in bags.

\begin{enumerate}[label=(\alph*)]

\item Draw a well-labeled graph of this market if Narnia does not allow trade. 
Calculate the equilibrium price and quantity, consumer surplus, producer surplus, 
and total surplus.

\item Narnia then opens the market to trade at the world price. Draw another graph 
to describe the new situation in the Turkish Delight market. Calculate the domestic 
price, quantities of consumption and production, imports, consumer surplus, 
producer surplus, total surplus, and the gains from trade. Using these values, 
determine who benefits and who loses from trade.

\item After a while, the White Witch imposes a \$1 per bag tariff on imported 
Turkish Delight in response to the pleas of Narnian sweet-makers. On a graph, 
show the effects of this tariff. Calculate the domestic price, quantities of 
consumption and production, imports, consumer surplus, producer surplus, 
government revenue, and total surplus.

\item What are the gains from opening up trade? What are the deadweight losses 
from restricting trade with the tariff? Who benefits under each policy? Who incurs a loss? Give both graphical and numerical answers.

\end{enumerate}

\section*{Question 2}

Parma is another land that is far, far away. Parma is ruled by King Giovanni, and the land both produces and consumes jawns. Parma's demand and supply of jawns are given by the following equations:
\begin{align*}
    Q_D &= 10 - \frac 1 2 P \\
    Q_S &= 2P
\end{align*}

where $P$ is measured in dollars per bag and $Q$ is measured in units.

\begin{enumerate}[label=(\alph*)]

\item Draw a well-labeled graph of this market if Parma does not allow trade. 
Calculate the equilibrium price and quantity, consumer surplus, producer surplus, 
and total surplus.

\item Parma then opens the market to international trade. The world price of jawns 
is $P_W = 8$. Draw another graph to describe the new situation in the Parmese 
jawn market. Calculate the domestic price, quantities of consumption and 
production, exports, consumer surplus, producer surplus, total surplus, 
and the gains from trade.

\item After hearing complaints from his dear friend Yinanni about the dearth of jawns,
King Giovanni imposes a \$2 per unit \textbf{export duty} on jawns leaving Parma. 
On a graph, show the effects of this export duty. Calculate the domestic price, 
quantities of consumption and production, exports, consumer surplus, 
producer surplus, government revenue, and total surplus.

\item What are the gains from opening up trade? What are the deadweight losses 
from restricting trade with the export duty? Who benefits under each policy? 
Who incurs a loss? Give both graphical and numerical answers.

\end{enumerate}
\end{document}
