\documentclass[12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load in packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{comment}
\usepackage[margin=1in]{geometry} % Sets 1-inch margins


%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{center}
\Large Week 1 practice problems

\medskip

\normalsize Elements of Microeconomics (discussion section 1)

\medskip

\small Rudy Kretschmar
\end{center}

\medskip

\subsection*{Question 1: Tradeoffs and Opportunity Cost}
You are trying to decide whether to go on a beach trip during spring break. What Tradeoffs would you need to consider in making this decision? What is the opportunity cost?

\subsection*{Question 2: Thinking on the Margin}
Speculate on whether or not you think the marginal cost or benefit of the specified item is large or small:

\begin{enumerate}
    \item The marginal benefit of your first glass of water for the day
    \item The marginal benefit of your seventeenth glass of water for the day 
    \item The cost of streaming an additional movie when you already have netflix
    \item The cost of streaming an additional movie on a streaming service you don't already own
    \item The marginal benefit of your first hour of studying for your upcoming econ 102 exam
    \item The marginal cost of your 30th hour studying in a row
    \item For an airline, the cost of flying one additional passenger on a flight that is not fully booked
    \item For an aireline, the cost of flying one additional passenger when the current flight is fully booked
    
\end{enumerate}

\subsection*{Question 3: More on Trade-offs}
What is the source of trade-offs? Can you give an example from your own life?

\begin{comment}
\subsection*{Question 2: PPF of a firm}
Let's start an Italian restaurant that makes pizzas and sandwiches.
\begin{enumerate}
    \item What will our production possibility frontier look like?
    \item Why will it take the shape that it has?
    \item How can we read the opportunity cost? Does it matter which part of the PPF we look at?
    \item Why might the shape change over time?
\end{enumerate}

\subsection*{Question 3: CPF of a consumer}
Now suppose we are \textit{going} to the Italian restaurant with a group of friends, and we want to decide what to order. Pizzas are \$10, sandwiches are \$5, and we have \$100 to spend.
    \begin{enumerate}
        \item What will our consumption possibility frontier look like?
        \item What is the opportunity cost of a pizza? Does it matter where on the CPF we are?
        \item What will happen to the CPF if we have \$200 to spend?
        \item What will happen to the CPF if the price of sandwiches increases to \$10?
    \end{enumerate}

\subsection*{Question 4: Italian restaurant economy}
Suppose the entire economy consists of Italian restaurants: during half of the week we work in one, and during the other half we buy food from them.
\begin{enumerate}
    \item What does the circular-flow diagram look like?
    \item What is missing from our model?
\end{enumerate}

\subsection*{Question 5: Guiding Economic Policy}
Economists are often asked to guide economic policy...
\medskip
\begin{enumerate}
    \item What are positive and normative statements?
    \begin{enumerate}
        \item Can you give an example of each?
    \end{enumerate}
    \item Why might two economists make different suggestions?
    \item Why might a politician ignore an economist's suggestions?
    
\end{enumerate}

\end{comment}

\end{document}